### 2025-02-17 manuscript.qmd
drafted methods

---
title: "Fine-scale predictions of the presence of predatory fish using a machine-learning framework"
author:
  - name: Theodore Hermann
    id: th
    orcid: 0000-0003-2632-7338
    email: ted.w.hermann@gmail.com
    affiliation:
    - name: Southwest Fisheries Science Center, Fisheries Ecology Division
      city: Santa Cruz
      state: CA
  - name: Peter Dudley
    id: pd
    email: peter.dudley@noaa.gov
    orcid: 0000-0002-3210-634X
    affiliation:
      - name: Southwest Fisheries Science Center, Fisheries Ecology Division
        city: Santa Cruz
        state: CA
abstract: >
  This is filler text.
  Test.
keywords:
  - salmonid
  - machine learning
  - Sacramento River
format:
  html: default
  docx: default
number-sections: true
bibliography: references.bib
csl: https://www.zotero.org/styles/ecological-modelling
---

```{r, echo=FALSE, message = FALSE, warning = FALSE}
library(targets)
```

# Introduction 
Freshwater biodiversity is threatened worldwide by anthropogenic activity [@allanOverfishingInlandWaters2005;@arthingtonFishConservationFreshwater2016;@reidEmergingThreatsPersistent2019]. Large, highly migratory fish spcies that require freshwater for all or part of their life histories are particularly vulnerable for many reasons, including diverse habitat requirements, large body size, and high market value [@heDisappearingGiantsReview2017;@stoneLastLeviathans2007;@waldmanNorthAmericanDiadromous2022]. Pacific salmonids (*Onchorhynchus* spp.) are culturally, economically, and ecologically important migratory fish that spawn in some of the most highly modified river systems in North America, which have been impacted by dams [@waplesEvolutionaryResponsesNative2008;@storchReviewPotentialConservation2022], agriculture [@macnealePesticidesAquaticFood2010], invasive species [@Kuehne2012;@Michel2018], channel modification [@richterEcologicallySustainableWater2003;@gregory29HistoricalChannel2007;@merzDisruptionNaturalDisturbance2024], and climate change [@siegelImpactsClimateChange2020].

The Sacramento River is the longest river in California and historically supported spawning and nursery ground for Pacific salmonids, providing an important food source for humans for thousands of years [@yoshiyamaHistoricalAbundanceDecline1998;@yoshiyamaHistorySalmonPeople1999]. However, the region has been impacted dramatically since the mid-nineteenth century by damming, mining, agriculture, fishing, and other anthropogenic activities [@munschOneHundredseventyYears2022]. Thus, the spawning migration runs of salmon have been greatly diminished and several are listed on the Endangered Species List [cite]. Habitat restoration projects have been undertaken for decades to support different aspects of salmonid life history in the Sacramento River basin [@goletSuccessesFailuresSuggested2013]. However, changing local habitat parameters can induce changes in populations of piscivorous fish that prey on juvenile salmonds [@Sabal2016] [any refs showing restoration unintetionally increasing preds??]. Juvenile salmonids are prey for both native predators, such as Sacramento pikeminnow (*Ptychocheilus grandis*), and numerous introduced predators, such as black bass (*Micropterus* spp.), striped bass (*Morone saxatilis*), white catfish (*Ameiurus catus*), and channel catfish (*Ictalurus punctatus*) [@Michel2018]. Understanding how restoration projects affect salmonid predators is thus crucial for management efforts.

Machine learning (ML) models are particularly suited to making new predictions from existing data and are growing in popularity for ecological applications [@rubbensMachineLearningMarine2023;@borowiecDeepLearningTool2022;@goodwinUnlockingPotentialDeep2022;@pichlerMachineLearningDeep2023]. Despite their effectiveness, many ML models are "black boxes," which reduces their interpretability and may undermine user confidence in the results compared with more familiar statistical models, especially when the results are used to make important decisions [@shinEffectsExplainabilityCausability2021;@rudinStopExplainingBlack2019]. However, several approaches exist for extracting meaning from ML models [@lucasTranslucentBoxInterpretable2020;@pichlerMachineLearningDeep2023], such as permutational importance [@altmannPermutationImportanceCorrected2010]. Another area of friction when using ML models in ecology is that there are many models to choose from, many potential applications, and no clear guidelines for which models may be appropriate. For example, sophisticated, nonlinear models, such as neural networks, may be most appropriate for some applications, but they are slow to train and difficult to interpret, and in some cases may offer little predictive advantage over a simpler model such as logistic regression [@pichlerMachineLearningDeep2023; @rudinOptimizedScoringSystems2018;@rudinStopExplainingBlack2019]. Thus, understanding and justifying the choice of model for a given application is important.

Here, we developed a robust ML pipeline for predicting the presence/absence of common predator of Sacramento River salmonids. To our knowledge, no study has explored the use of machine learning models to predict the presence/absence of predatory freshwater fish at a fine spatial scale based on environmental parameters. Our pipeline includes several linear and nonlinear models to explore which algorithms may be most appropriate for this application. Models are compared based on predictive performance and training time [!!TODO!!]. We demonstrate the use of the pipeline with two case studies (smallmouth bass and Sacramento pikeminnow from the lower Sacremento River and largemouth bass from the Sacramento River delta) based on field data. The framework outlined here is flexible enough that it could be applied to other river systems where similar data are available.

# Methods (do we need a "study area" section?)
<!-- 
## study areas
 - Sacramento River
 The American River is one of the major tributaries of California’s Sacramento River. Historically, the American River supported spring-, fall-, and late-fall-run salmon, but gold placer mining and later upstream dam construction have reduced the runs to a fall run. For juveniles in the river, suitable habitat exists only within the Lower American River downstream of Nimbus Dam. The American River has been subject of management and restoration efforts (Williams Bulletin) in attempts to increase the use of salmon spawning sites (Zeug et al., 2014), increase adult salmon production (Peterson and Duarte 2020), and enhance juvenile salmon rearing habitat (Hellmair et al., 2018).

 - Sacramento River delta
 - Do I need a map here?? -->

## Data sources
### Sacramento River
Models for the lower Sacramento were trained and tested using field data collected in 2013 and 2014 [@FISHBIO2014; @FISHBIO2015a]. Briefly, transects starting 15 feet from shore and moving towards shore were electroshocked and all fish were collected. Afterwards, environmental data were collected from each transect, including cover from submerged vegetation and woody debris (percent), presence/absence of shade, type of substrate, water depth (at 5 ft, 10 ft, and 15 ft from shore) and water velocity (at 5 ft, 10 ft, and 15 ft from shore). The data were then cleaned and parsed. Vegetation and wood data were originally reported as percent ranges. The mean value of each range was used and converted to a proportion. Substrate classification varied among study years, and the vast majority of values were either rocky or not; therefore, all substrate classifications that included the word "rocky" were simplified to "rock". Gravel, which appeared very infrequently, was classified as "rock." All other substrates were classified as "not rock". Finally, water depth and velocity values were averaged for each transect (i.e., the respective values at 5, 10, and 15 ft from shore) to yield a single depth or velocity value per transect.

Although the data set provided records of several piscivorous fish species, we only used data for  smallmouth bass (*Micropterus dolomieu*) and Sacramento pikeminnow (*Ptychocheilus grandis*). Predators such as striped bass (*Morone saxatilis*) and channel catfish (*Ictalurus punctatus*), which feed on juvenile fish in the Sacramento Delta and can have a large impact on salmonids [e.g., @Lindley2003; @Michel2018], were excluded because only a individuals were recorded over a several year period. This indicates either that these species are not a significant influence on juvenile salmonid and sturgeon populations in this part of the river, or that they were simply missed during sampling. Regardless, the data were inadequate for modeling and therefore these species were excluded from this analysis.

Data from 2013 reported only observations of *Micropterus* species (listed as “black bass”), whereas data in 2014 reported *Micropterus* observations by species. For consistency with the 2013 observations, all black bass data from 2014 were pooled. Hereafter, references to “smallmouth bass” refer to pooled “black bass” data. Furthermore, observations included fish size (recorded as < 150 mm and > 150 mm). Due to the limited numbers of total observations in the data, we pooled fish of all sizes per species, even though smaller fish are less likely to be piscivorous and may exhibit different behaviors and habitat choices compared with larger conspecifics. Finally, the original count data were transformed into presence/absence data because there were so few transects in which more than one individual per predator species was observed.

### Sacramento--San Joaquin Delta -- how to cite?
Our second data set was collected in the Sacramento--San Joaquin Delta as part of the California Department of Fish and Game’s (CDFG) Delta Resident Shoreline Fish Monitoring Project. This survey was conducted sporadically in three decades (1980s, 1990s, and 2000s) until December 2004. Shoreline areas were shocked for about 10 to 15 seconds in 20 m increments. Data were collected from each sampling area including location information, water quality (temperature, water clarity [Secchi disc, cm], conductivity [µmhos], and turbidity [NTU]), habitat characteristics (channel type, sampling area, vegetation type, bank type, depth [feet], and number of snags), environmental variables (surface flow (water velocity [m/s]), estimated wind speed [mph], and light [LUX]), electro-fishing settings (including time and distance of shoreline shocked), and catch (fish species count and length). We focused on largemouth bass since they were abundant in the data and a previous study using this data set found that largemouth bass had a high detection probability using the these sampling methods [@mckenzieEvaluatingRoleBoat2021], suggesting a lower occurrence of false negatives in the data. For consistency with the data from the Sacramento River, we converted counts into presencen/absence.

## Models
We included three linear models in the study: a general linear model (GLM) without interaction or polynomial terms, was used as a baseline comparison for other models in the study; a GLM with interactions and/or polynomial terms (several combinations used in the workflow); and a GLM with L1 and L2 regularization ("elastic net") [@zouRegularizationVariableSelection2005]. We also included four nonlinear models: a support vector machine (SVM) with radial basis function kernel, which optimizes a hyperplane separating categories [@scholkopfComparingSupportVector1997]; a random forest classifier, which is an ensemble of decision trees that grows randomly with subsampled data [@biauRandomForestGuided2016]; a gradient-boosted tree classifier (XGBoost), which is an ensemble of decision trees that grows greedily [@chenXGBoostScalableTree2016]; a multilayer perceptron model, which is a single layer, feed-forward neural network [@hornikMultilayerFeedforwardNetworks1989]; and an bagged neural net, which is a collection of neural networks forming an ensemble via bootstrap aggregating ("bagging") [@breimanBaggingPredictors1996].

## Training and evaluation
Models were trained using the `tidymodels` metapackage (v. `r packageVersion("tidymodels")`) in R (v. `r getRversion()`). Preprocessing of data differed for each model and followed @kuhnFeatureEngineeringSelection2019. Because there were many more absences in the data, we upsampled rows with presences in the training data using the R package `themis` (v. `r packageVersion("themis")`). Otherwise, the optimal model may simply have predicted only absences. Models required hyperparameter tuning during training, which varied with the model. The base GLM and GLM with interactions and/or polynomial terms had no hyperparameters to tune. For the elastic net GLM, we tuned the penalty and mixture parameters, which determined the total amount of regularization and the mix of L1 and L2 regularization, respectively. For the SVM, we tuned cost and sigma, which influence the strength of regularization and decision boundary, respectively. For the random forest, we tuned the number of predictors randomly sampled at each split during training, the number of trees, and the minimum number of data points in a node required for the node to be split further. For XGBoost, we tuned the same parameters as those for the random forest model in addition to the maximum depth of trees, learning rate, the fraction of data exposed to the fitting routine, and the reduction in the loss function required for further splits. For the two neural network models, we tuned the number of units in the hidden model, the weight decay, and the number of training iterations.

Given the computational requirements to train this many models, we took several steps to improve computation time. First, we parallelized our workflow were possible using the R packages `crew` (v. `r packageVersion("crew")`) and `targets` (v. `r packageVersion("targets")`). Additionally, we used a Latin hypercube grid search [@husslageSpaceFillingLatinHypercube2008] rather than a granular grid search to determine the search space for hyperparameters. This allowed searching a nearly random sample of parameter values without requiring an exhaustive search over a much larger granular combination of values. Finally, we performed our grid search with an ANOVA race with the R package `finetune` (v. `r packageVersion("finetune")`) [@kuhnFutilityAnalysisCrossValidation2014]. Briefly, after a burn in period of 3 resamples, the performance of hyperparameter combinations was evaluated at each resampling step using repeated-measures ANOVA to eliminate those that were unlikely to be the best. Thus, the best performing model can often be identified without performing all specified resampling steps. Even if all resampling steps are performed, the number of models evaluated at each step is usually reduced, improving overall computation time.

<!-- ### pipeline - Do we need a flow diagram?
- data split: 80% training data, 20% reserved for testing data
- training data used for X-fold cross validation repeated x times.
- Where appropriate, models were trained using different hyperparameter combinations.
- hyperparameters were selected using a latin hypercube to provide good coverage of values without needing to use an exhaustive full grid search approach
- models trained using tidymodels framework
- each model tested with various feature engineering options following xxxx
- the importance of model features was estimated using permutation;
- permutation importance is a model-agnostic tool that can be used to compare feature importance across different models -->

## Data analysis
All analyses and visualizations were conducted in R. The importance of model features was estimated with permutation importance using the R package `vip` (v. `r packageVersion("vip")`). Figures were made using `ggplot2` (v. `r packageVersion("ggplot2")`).

## Code
All code is available on Github, including the code for producing this manuscript: [https://github.com/mrguyperson/pred_predictive_modeling](https://github.com/mrguyperson/pred_predictive_modeling).

# Results
## Modeling pipeline

![roc auc plot](`r tar_read(roc_plot)`){#fig-roc-auc}

![var imp plot](`r tar_read(var_imp_plot)`){#fig-var-imp}

![histogram](/output/predator_histograms.jpg){#fig-histogram}

![maps](/output/hab_rating_all_models.jpg){#fig-maps}

## model selection

## model performance

## model interpretability

# Discussion

## trade-offs among models
- nonlinear models can be harder to interpret but may offer improved predictions
- some models take considerably longer to train than others but may not offer considerable advantages for predictions.

## method may be species specific
- hard to collect data for species that appear sporadically in large numbers
- in effective for species that are highly mobile

## limitations for "cell" size when collecting data
- sampling area size should reflect the behavior of a fish
- if data transects are too large, then environmental heterogeneity may be overly "averaged"; i.e., each transect may be too similar
- similarly, larger areas are more likely to contain fish, so the model may simply predict fish in every "cell"
- areas that are too small may inflate the number of absences in the data, so the model may simply predict that all cells are empty.

## data availability limitations
- data of this nature are time and energy intensive to collect, especially if they are to include the adequate spatiotemporal coverage
- data are also improved if sampling is exhaustive; e.g., multiple passes with electroshockers to catch as many fish as possible.